---
title: "Assignment9"
author: "Luke Van Horn"
date: "1/30/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---
1. Using a for loop and an if statement, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a vector of integers with some zero values. Then, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.
```{r}

v <- sample(rep(0:3, each=7))
v
counter<-0


for (i in 1:length(v)){
  if (v[i]==0) {
  counter <- counter + 1
  }
}
counter
```


2. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}


f<-function(x, y){
  d<-matrix(NA, nrow=x, ncol=y)
  for (i in 1:x){
        for (j in 1:y){
          d[i,j] <-i*j
        }
  }
 return(d)
}

f(5,6)

```

3. Use a for loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows).
```{r}
funky<-function(p=NULL){
  if(is.null(p)) {
    Type<-c(rep("Control", 25), rep("Treatment", 25))
    Value<-c(rnorm(25, mean=10, sd=1.5),rnorm(25, mean=45, sd=2))
    p<-data.frame(Type,Value)
    aovfunk<-aov(Value~Type, data=p)
    pval<-summary(aovfunk)[[1]][["Pr(>F)"]][1]
  }
  return(pval)
}
funk<-funky()
funk

df<-data.frame()
for(i in 1:100){
  df[i,1] <-i
  df[i,2] <- funky()
}
df

# attempt in one function
fresh<-function(p=NULL){
    if(is.null(p)) {
  df<-data.frame()
    for (i in 1:100){
         luke<- function(p=NULL){
                    if(is.null(p)) {
                Type<-c(rep("Control", 25), rep("Treatment", 25))
                Value<-c(rnorm(25, mean=10, sd=1.5),rnorm(25, mean=45, sd=2))
                p<-data.frame(Type,Value)
                aovfunk<-aov(Value~Type, data=p)
                pval<-summary(aovfunk)[[1]][["Pr(>F)"]][1]
                    }
             return(pval)
         }
         df[i,1] <- luke()
    }
    }
  return(df)
}
fresh()



```



