---
title: "FinalProject"
author: "Luke Van Horn"
date: "1/31/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---
## About `leaflet`  
`leaflet` is an open-source JavaScript library that allows creation of interactive maps. I chose to learn about `leaflet` because of its aesthetic maps and coding ease. Having practice with a mapping package rounds out my R capabilities nicely.  
## Leaflet basics

In order to get started with leaflet, I downloaded `leaflet` by running `install.packages("leaflet") in my console. To get familiar with the package, I made a basic map called "map" with the default OpenStreetMap map tiles. This map creates a labelled popup called "Middlebury College" on the longitute and latitute coordinates I found online for Middlebury College. 
```{r}
library("leaflet")
map <- leaflet()%>%
  addTiles() %>%
  addMarkers(lng=-73.17594474252131, lat=44.00826191138904, popup="Middlebury College")
map
```
Data can be taken from data frames and put onto interactive maps. Below is some code that maps 10 points with longitute and latitude values.
```{r}
df = data.frame(Lat = 1:10, Long = rnorm(10))
df
leaflet(df) %>% addCircles(lng = ~Long, lat = ~Lat)
```
Although leaflet will automatically create circles from "Long" and "Lat" columns with `addCircles()`, this can explicitly stated with `lng = ~Long, lat = ~Lat`.

```{r}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m = leaflet(df) %>% addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)
m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))

```


